Visual Diagnostics for Constrained Optimisation with Application to Guided Tours

### 2021-01-28 paper/zhang-cook-laa-langrene-menendez.Rmd
text editing on search, trace and pca plot WIP

\begin{algorithm}
\SetAlgoLined
  \SetKwInOut{input}{input}
  \SetKwInOut{output}{output}
    \input{$f(.)$, $\alpha_1$, $l_{\max}$, $\text{cooling}$} 
    \output{$\mathbf{A}_{l}$}
    generate random start $\mathbf{A}_1$ and set $\mathbf{A}_{\text{cur}} \coloneqq \mathbf{A}_1$, $I_{\text{cur}} = f(\mathbf{A}_{\text{cur}})$, $j = 1$\;
  \Repeat{$\mathbf{A}_l$ is too close to $\mathbf{A}_{\text{cur}}$ in terms of geodesic distance}{
   set $l = 1$\;
  \Repeat{$l > l_{\max}$ or $I_{l} > I_{\text{cur}}$}{
    generate $\mathbf{A}_{l} = (1- \alpha_j)\mathbf{A}_{\text{cur}} + \alpha_j \mathbf{A}_{\text{rand}}$ and orthogonalise $\mathbf{A}_{l}$\;
    compute $I_{l}  = f(\mathbf{A}_{l})$\;
    update $l = l + 1$\;
  }
  update $\alpha_{j+1} = \alpha_j * \text{cooling}$\;
  construct the geodesic interpolation between $\mathbf{A}_{\text{cur}}$ and $\mathbf{A}_l$\; 
  update $\mathbf{A}_{\text{cur}} = \mathbf{A}_l$ and $j = j + 1$\;
}
  \caption{random search}
  \label{random-search}
\end{algorithm}

@@

\begin{algorithm}
\SetAlgoLined
\Repeat{$l > l_{\max}$ or $I_{l} > I_{\text{cur}}$ or $P > U$}{
    generate $\mathbf{A}_{l} = (1- \alpha_j)\mathbf{A}_{\text{cur}} + \alpha_j \mathbf{A}_{\text{rand}}$ and orthogonalise $\mathbf{A}_{l}$\;
    compute $I_{l}  = f(\mathbf{A}_{l})$, $T(l) = \frac{T_0}{\log(l + 1)}$ and $P= \min\left\{\exp\left[-\frac{I_{\text{cur}} -I_{l}}{T(l)}\right],1\right\}$\;
    draw $U$ from a uniform distribution: $U \sim \text{Unif(0, 1)}$\;
    update $l = l + 1$\;
  }
  \caption{simulated annealing}
  \label{simulated-annealing}
\end{algorithm}

@@

\begin{algorithm}
\SetAlgoLined
\Repeat{$l > l_{\max}$ or $p_{\text{diff}} > 0.001$}{
  generate $n$ random directions $\mathbf{A}_{\text{rand}}$ \;
  compute $2n$ candidate bases deviate from $\mathbf{A}_{\text{cur}}$ by an angle of $\delta$ while ensure orthogonality\;
  compute the corresponding index value for each candidate bases\;
  determine the search direction as from $\mathbf{A}_{\text{cur}}$ to the candidate bases with the largest index value\;
  determine the step size via optimising the index value on the search direction over a 90 degree window\;
  find the optima $\mathbf{A}_{**}$ and compute $I_{**} = f(\mathbf{A}_{**})$, $p_{\text{diff}} = (I_{**} - I_{\text{cur}})/I_{**}$\;
  update $l = l + 1$\;
}
\caption{search geodesic}
\label{search-geodesic}
\end{algorithm}

In pseudo derivative [@cook1995grand] method, the search direction is computed using the most prominent direction that deviating an tiny angle of $\delta$ from the current basis. The step size is chosen by optimising the index value along the geodesic direction over an 90 degree angle from $-\pi/4$ to $\pi/4$ along the search direction chosen. The optima $\mathbf{A}_{**}$ is returned for the current iteration if it meets the percentage improve condition or when $l_{\max}$ is reached.  Algorithm \ref{search-geodesic} summarises the inner loop of the pseudo derivative search.

@@

A primary interest of diagnosing an optimiser is to study how it progressively finds its optimum. A simple treatment of plotting the index value across its natural order will cause the graph to be disproportional to the iteration since it usually takes much longer for an optimiser to find a better basis towards the end. Another option is to use summarisation in each iteration. Boxplot is a suitable candidate that can provide five points summary of each iteration and additional information can be separately added with new layers, for example, text information on the number of points searched in each iteration is added at the bottom of each iteration, the bases returned for interpolation are highlighted in larger size and linked, and information regarding the last iteration where no basis is returned is turned to grey scale. Furthermore, an option to switch from boxplot back to point geometry is helpful when the number of observation is small in one iteration and can be achieved via the `cutoff` argument.

Figure \ref{fig:toy-search} shows the searches of two different optimisers: SA (left) and SAJO (right). Both optimisers progress quickly at the first few iterations, take longer to find better basis in the later iterations, and finish off when hitting the maximum number of try, 499 and 999, respectively. Also, the target basis found on the left panel always has an increased index value in each iteration while this is not the case for the plot on the right. This explains why in this example, the optimiser SAJO takes longer to find the final basis, but in more complicated scenarios, the feature of probabilistic acceptance allows a more holistic search of the basis space and is more likely to find the global optimum.

```{r toy-search, fig.env = "widefigure", fig.cap="A comparison of search by two otimisers: SA and SAJO on a six-variable dataset \\code{boa6} with holes index on a 2D problem. Both optimisers finish when the maximum number of try is reached and attain similar index value in the end. Simulated annealing with jump out (SAJO) takes longer to get to the final basis because it allows for probabilistic acceptance of inferior basis, which is a feature useful for more complex problem."}
```

@@

```{r toy-interp, fig.env = "widefigure", fig.cap = "Trace plots of the bases on the interpolation path for two optimisers with the same configuration as the previous figure. The right path shows that to reach a lower target basis, the interpolation actually first passes some higher basis and this could potentially be information to be incorporated into the optimisers."}
```

Viewing the index value of points on the interpolation path is another interest to the analysts since the projection on these bases will be played by the tour animation and it provides further information on how the index value changes when moving from one target basis to another. Figure \ref{fig:toy-interp} presents the index value against time on the interpolation path for the two optimisers with same configuration as \ref{fig:toy-search}. From Figure \ref{fig:toy-interp}, we further know that to reach a target basis with lower index value, the optimiser plotted on the right  first passes some bases with higher index value during the interpolation. If the bases on the interpolation path can be incorporated into the optimisation, we may be able to find the optimal basis quicker.

@@

```{r toy-pca, fig.height=3, fig.cap = "1D projection on the 5-variable dataset \\code{boa5} with two optimisers: Simulated Annealing (SA) and Pseudo Derivative (PD). All the bases in pseudo derivative has been flipped positive so that the two paths finish close to each other and to the theoretical best."}
```

Apart from checking the progression of an optimiser, another interesting aspect of the diagnostics is to visualise how the visited bases looks like in its parameter space. Given the orthonormality constraint, the space of projection bases $\mathbf{A}_{p \times d}$ is a $p \times d$ dimension sphere and even with 1D projection, the basis space is high-dimensional. Dimension reduction methods, i.e. principal component analysis can be used to present the bases on 2D space with appropriate annotation to highlight the different key components of the optimisation.

In a projection pursuit guided tour optimisation, there are 7 different components that are worth-noticing in the exploration: 1) The spherical space of all the basses; 2) The starting basis of an optimisation routine; 3) The anchor points where a new target basis is found; 4) The search basis that an optimiser has evaluated; 5) The interpolation path that constructed by the guided tour; 6) The annotation of the interrupted path and 7) The theoretical best basis, if applicable. Despite all have to be displayed on the diagnostic plot, these components differ in their importance with the most interesting ones being the starting point, anchor points and the interpolation path.

>>>>

\begin{algorithm}
\SetAlgoLined
  \SetKwInOut{input}{input}
  \SetKwInOut{output}{output}
    \input{$f(.)$, $\alpha_1$, $l_{\max}$, $\text{cooling}$} 
    \output{$\mathbf{A}_{l}$}
    generate random start $\mathbf{A}_1$ and set $\mathbf{A}_{\text{cur}} \coloneqq \mathbf{A}_1$, $I_{\text{cur}} = f(\mathbf{A}_{\text{cur}})$, $j = 1$\;
  \Repeat{$\mathbf{A}_l$ is too close to $\mathbf{A}_{\text{cur}}$ in terms of geodesic distance}{
   set $l = 1$\;
  \Repeat{$l > l_{\max}$ or $I_{l} > I_{\text{cur}}$}{
    generate $\mathbf{A}_{l} = (1- \alpha_j)\mathbf{A}_{\text{cur}} + \alpha_j \mathbf{A}_{\text{rand}}$ and orthogonalise $\mathbf{A}_{l}$\;
    compute $I_{l}  = f(\mathbf{A}_{l})$\;
    update $l = l + 1$\;
  }
  update $\alpha_{j+1} = \alpha_j * \text{cooling}$\;
  construct the geodesic interpolation between $\mathbf{A}_{\text{cur}}$ and $\mathbf{A}_l$\; 
  update $\mathbf{A}_{\text{cur}} = \mathbf{A}_l$ and $j = j + 1$\;
}
  \caption{Simulated annealing (SA)}
  \label{random-search}
\end{algorithm}

@@

\begin{algorithm}
\SetAlgoLined
\Repeat{$l > l_{\max}$ or $I_{l} > I_{\text{cur}}$ or $P > U$}{
    generate $\mathbf{A}_{l} = (1- \alpha_j)\mathbf{A}_{\text{cur}} + \alpha_j \mathbf{A}_{\text{rand}}$ and orthogonalise $\mathbf{A}_{l}$\;
    compute $I_{l}  = f(\mathbf{A}_{l})$, $T(l) = \frac{T_0}{\log(l + 1)}$ and $P= \min\left\{\exp\left[-\frac{I_{\text{cur}} -I_{l}}{T(l)}\right],1\right\}$\;
    draw $U$ from a uniform distribution: $U \sim \text{Unif(0, 1)}$\;
    update $l = l + 1$\;
  }
  \caption{Simulated annealing with jump out (SAJO)}
  \label{simulated-annealing}
\end{algorithm}

@@

\begin{algorithm}
\SetAlgoLined
\Repeat{$l > l_{\max}$ or $p_{\text{diff}} > 0.001$}{
  generate $n$ random directions $\mathbf{A}_{\text{rand}}$ \;
  compute $2n$ candidate bases deviate from $\mathbf{A}_{\text{cur}}$ by an angle of $\delta$ while ensure orthogonality\;
  compute the corresponding index value for each candidate bases\;
  determine the search direction as from $\mathbf{A}_{\text{cur}}$ to the candidate bases with the largest index value\;
  determine the step size via optimising the index value on the search direction over a 90 degree window\;
  find the optima $\mathbf{A}_{**}$ and compute $I_{**} = f(\mathbf{A}_{**})$, $p_{\text{diff}} = (I_{**} - I_{\text{cur}})/I_{**}$\;
  update $l = l + 1$\;
}
\caption{Pseudo derivative (PD)}
\label{search-geodesic}
\end{algorithm}

In pseudo derivative search [@cook1995grand], the search direction is computed using the most prominent direction that deviating an tiny angle of $\delta$ from the current basis. The step size is chosen by optimising the index value along the geodesic direction over an 90 degree angle from $-\pi/4$ to $\pi/4$ along the search direction chosen. The optima $\mathbf{A}_{**}$ is returned for the current iteration if it meets the percentage improve condition or when $l_{\max}$ is reached.  Algorithm \ref{search-geodesic} summarises the inner loop of the pseudo derivative search.

@@

A starting point of diagnosing an optimiser is to understand how many search the optimiser has conducted. One may want to simply plot the index value of the search points across its natural order. A point geometry may work well if each iteration has a similar number of points,  however, the plot will be disproportionate if some iterations have considerably more points than other - it over-emphasizes the iterations that have more search points! An alternative is to summarise the search in each iteration using boxplot where each iteration will be spaced out equally. Occasionally, one may still want to switch back to point geometry if the number of points is small and this can be achieved via the `cutoff` argument. Additional annotation can be added with new layers, for example, the number of points searched in each iteration can be added as text label at the bottom of each iteration; the anchor bases to interpolate are connected and highlighted in a larger size; and the color of the last iteration is in a grey scale to indicate no better basis found in this iteration.

Figure \ref{fig:toy-search} shows the searches of two different optimisers: SA (left) and SAJO (right). Both optimisers quickly find better bases at the first few iterations,  then take longer in the later iterations, and finally finish when the maximum number of evaluation is reached. The anchor bases, the ones with the highest index value in each iteration, always have an increased index value in the optimiser SA while this is not the case for SAJO. This feature gives SA an advantage in simple example, like the one presented while the jump out component in SAJO allows for a more holistic search in more complicated examples.

```{r toy-search, fig.env = "widefigure", fig.cap="A comparison of search by two otimisers: SA (left) and SAJO (right) on a 2D projection of a six-variable dataset, \\code{boa6}, using holes index. Both optimisers find the final basis with similar index value while it takes SAJO more iteration because the algorithm allows a probabilistic acceptance of bases with lower index value, as observed in iteration 4 and 6-8."}
```

@@

```{r toy-interp, fig.env = "widefigure", fig.cap = "Trace plots of the interpolating bases with the same optimisation routine as the previous figure. Both traces are smooth while the change of index value in each iteration may not be monotinic."}
```

Another interest of diagnosing is to examine how the index value progresses for the points on the interpolation path since the projection on these bases are played by the tour animation. Trace plots are created with plotting the index value of the interpolation points against time. Figure \ref{fig:toy-interp} presents the trace plot of the two optimisation routines as \ref{fig:toy-search}. The trace plot is smooth in both plots and this is a character of the holes index function used. An interesting discovery is that with optimiser SAJO, the interpolator, in iteration 6-8, first passes through some bases with higher index value than the anchor bases.

@@

```{r toy-pca, fig.height=3, fig.cap = "1D projection on the 5-variable dataset \\code{boa5}5) with two optimisers: Simulated Annealing (SA) and Pseudo Derivative (PD). All the bases in pseudo derivative has been flipped positive so that the two paths finish close to each other and to the theoretical best."}
```

Apart from checking the progression of an optimiser, looking at how the points evaluated by the optimisers look like in the space is another interest. Given the orthonormality constraint, the space of projection bases $\mathbf{A}_{p \times d}$ is a $p \times d$ dimension sphere and  dimension reduction methods, i.e. principal component analysis is applied first to project all the bases onto a 2D space. In a projection pursuit guided tour optimisation, there are various type of bases involved: 1) The starting basis; 2) The anchor bases that have the highest index value in each iteration; 3) The search bases that an optimiser evaluated to produce the anchor bases; 4) The interpolating bases on the interpolation path; and finally 5) the ending basis. The importance of these basis differs with the most important ones being the start, interpolating and ending bases. Anchor and search bases can be turned on with argument `details = TRUE`.

Several annotations have been made to facilitate better reading of this plot.

### 2021-01-28 paper/zhang-cook-laa-langrene-menendez.tex
text editing on search, trace and pca plot WIP

\begin{algorithm}
\SetAlgoLined
\Repeat{$l > l_{\max}$ or $p_{\text{diff}} > 0.001$}{
  generate $n$ random directions $\mathbf{A}_{\text{rand}}$ \;
  compute $2n$ candidate bases deviate from $\mathbf{A}_{\text{cur}}$ by an angle of $\delta$ while ensure orthogonality\;
  compute the corresponding index value for each candidate bases\;
  determine the search direction as from $\mathbf{A}_{\text{cur}}$ to the candidate bases with the largest index value\;
  determine the step size via optimising the index value on the search direction over a 90 degree window\;
  find the optima $\mathbf{A}_{**}$ and compute $I_{**} = f(\mathbf{A}_{**})$, $p_{\text{diff}} = (I_{**} - I_{\text{cur}})/I_{**}$\;
  update $l = l + 1$\;
}
\caption{search geodesic}
\label{search-geodesic}
\end{algorithm}

In pseudo derivative \citep{cook1995grand} method, the search direction
is computed using the most prominent direction that deviating an tiny
angle of \(\delta\) from the current basis. The step size is chosen by
optimising the index value along the geodesic direction over an 90
degree angle from \(-\pi/4\) to \(\pi/4\) along the search direction
chosen. The optima \(\mathbf{A}_{**}\) is returned for the current
iteration if it meets the percentage improve condition or when
\(l_{\max}\) is reached. Algorithm \ref{search-geodesic} summarises the
inner loop of the pseudo derivative search.

@@

A primary interest of diagnosing an optimiser is to study how it
progressively finds its optimum. A simple treatment of plotting the
index value across its natural order will cause the graph to be
disproportional to the iteration since it usually takes much longer for
an optimiser to find a better basis towards the end. Another option is
to use summarisation in each iteration. Boxplot is a suitable candidate
that can provide five points summary of each iteration and additional
information can be separately added with new layers, for example, text
information on the number of points searched in each iteration is added
at the bottom of each iteration, the bases returned for interpolation
are highlighted in larger size and linked, and information regarding the
last iteration where no basis is returned is turned to grey scale.
Furthermore, an option to switch from boxplot back to point geometry is
helpful when the number of observation is small in one iteration and can
be achieved via the \texttt{cutoff} argument.

Figure \ref{fig:toy-search} shows the searches of two different
optimisers: SA (left) and SAJO (right). Both optimisers progress quickly
at the first few iterations, take longer to find better basis in the
later iterations, and finish off when hitting the maximum number of try,
499 and 999, respectively. Also, the target basis found on the left
panel always has an increased index value in each iteration while this
is not the case for the plot on the right. This explains why in this
example, the optimiser SAJO takes longer to find the final basis, but in
more complicated scenarios, the feature of probabilistic acceptance
allows a more holistic search of the basis space and is more likely to
find the global optimum.

@@

\caption[A comparison of search by two otimisers]{A comparison of search by two otimisers: SA and SAJO on a six-variable dataset \code{boa6} with holes index on a 2D problem. Both optimisers finish when the maximum number of try is reached and attain similar index value in the end. Simulated annealing with jump out (SAJO) takes longer to get to the final basis because it allows for probabilistic acceptance of inferior basis, which is a feature useful for more complex problem.}\label{fig:toy-search}
\end{widefigure}
\end{Schunk}

@@

\caption[Trace plots of the bases on the interpolation path for two optimisers with the same configuration as the previous figure]{Trace plots of the bases on the interpolation path for two optimisers with the same configuration as the previous figure. The right path shows that to reach a lower target basis, the interpolation actually first passes some higher basis and this could potentially be information to be incorporated into the optimisers.}\label{fig:toy-interp}
\end{widefigure}
\end{Schunk}

Viewing the index value of points on the interpolation path is another
interest to the analysts since the projection on these bases will be
played by the tour animation and it provides further information on how
the index value changes when moving from one target basis to another.
Figure \ref{fig:toy-interp} presents the index value against time on the
interpolation path for the two optimisers with same configuration as
\ref{fig:toy-search}. From Figure \ref{fig:toy-interp}, we further know
that to reach a target basis with lower index value, the optimiser
plotted on the right first passes some bases with higher index value
during the interpolation. If the bases on the interpolation path can be
incorporated into the optimisation, we may be able to find the optimal
basis quicker.

@@

\caption[1D projection on the 5-variable dataset \code{boa5} with two optimisers]{1D projection on the 5-variable dataset \code{boa5} with two optimisers: Simulated Annealing (SA) and Pseudo Derivative (PD). All the bases in pseudo derivative has been flipped positive so that the two paths finish close to each other and to the theoretical best.}\label{fig:toy-pca}
\end{figure}
\end{Schunk}

Apart from checking the progression of an optimiser, another interesting
aspect of the diagnostics is to visualise how the visited bases looks
like in its parameter space. Given the orthonormality constraint, the
space of projection bases \(\mathbf{A}_{p \times d}\) is a
\(p \times d\) dimension sphere and even with 1D projection, the basis
space is high-dimensional. Dimension reduction methods, i.e.~principal
component analysis can be used to present the bases on 2D space with
appropriate annotation to highlight the different key components of the
optimisation.

In a projection pursuit guided tour optimisation, there are 7 different
components that are worth-noticing in the exploration: 1) The spherical
space of all the basses; 2) The starting basis of an optimisation
routine; 3) The anchor points where a new target basis is found; 4) The
search basis that an optimiser has evaluated; 5) The interpolation path
that constructed by the guided tour; 6) The annotation of the
interrupted path and 7) The theoretical best basis, if applicable.
Despite all have to be displayed on the diagnostic plot, these
components differ in their importance with the most interesting ones
being the starting point, anchor points and the interpolation path.

>>>>

\begin{algorithm}
\SetAlgoLined
\Repeat{$l > l_{\max}$ or $p_{\text{diff}} > 0.001$}{
  generate $n$ random directions $\mathbf{A}_{\text{rand}}$ \;
  compute $2n$ candidate bases deviate from $\mathbf{A}_{\text{cur}}$ by an angle of $\delta$ while ensure orthogonality\;
  compute the corresponding index value for each candidate bases\;
  determine the search direction as from $\mathbf{A}_{\text{cur}}$ to the candidate bases with the largest index value\;
  determine the step size via optimising the index value on the search direction over a 90 degree window\;
  find the optima $\mathbf{A}_{**}$ and compute $I_{**} = f(\mathbf{A}_{**})$, $p_{\text{diff}} = (I_{**} - I_{\text{cur}})/I_{**}$\;
  update $l = l + 1$\;
}
\caption{Pseudo derivative (PD)}
\label{search-geodesic}
\end{algorithm}

In pseudo derivative search \citep{cook1995grand}, the search direction
is computed using the most prominent direction that deviating an tiny
angle of \(\delta\) from the current basis. The step size is chosen by
optimising the index value along the geodesic direction over an 90
degree angle from \(-\pi/4\) to \(\pi/4\) along the search direction
chosen. The optima \(\mathbf{A}_{**}\) is returned for the current
iteration if it meets the percentage improve condition or when
\(l_{\max}\) is reached. Algorithm \ref{search-geodesic} summarises the
inner loop of the pseudo derivative search.

@@

A starting point of diagnosing an optimiser is to understand how many
search the optimiser has conducted. One may want to simply plot the
index value of the search points across its natural order. A point
geometry may work well if each iteration has a similar number of points,
however, the plot will be disproportionate if some iterations have
considerably more points than other - it over-emphasizes the iterations
that have more search points! An alternative is to summarise the search
in each iteration using boxplot where each iteration will be spaced out
equally. Occasionally, one may still want to switch back to point
geometry if the number of points is small and this can be achieved via
the \texttt{cutoff} argument. Additional annotation can be added with
new layers, for example, the number of points searched in each iteration
can be added as text label at the bottom of each iteration; the anchor
bases to interpolate are connected and highlighted in a larger size; and
the color of the last iteration is in a grey scale to indicate no better
basis found in this iteration.

Figure \ref{fig:toy-search} shows the searches of two different
optimisers: SA (left) and SAJO (right). Both optimisers quickly find
better bases at the first few iterations, then take longer in the later
iterations, and finally finish when the maximum number of evaluation is
reached. The anchor bases, the ones with the highest index value in each
iteration, always have an increased index value in the optimiser SA
while this is not the case for SAJO. This feature gives SA an advantage
in simple example, like the one presented while the jump out component
in SAJO allows for a more holistic search in more complicated examples.

@@

\caption[A comparison of search by two otimisers]{A comparison of search by two otimisers: SA (left) and SAJO (right) on a 2D projection of a six-variable dataset, \code{boa6}, using holes index. Both optimisers find the final basis with similar index value while it takes SAJO more iteration because the algorithm allows a probabilistic acceptance of bases with lower index value, as observed in iteration 4 and 6-8.}\label{fig:toy-search}
\end{widefigure}
\end{Schunk}

@@

\caption[Trace plots of the interpolating bases with the same optimisation routine as the previous figure]{Trace plots of the interpolating bases with the same optimisation routine as the previous figure. Both traces are smooth while the change of index value in each iteration may not be monotinic.}\label{fig:toy-interp}
\end{widefigure}
\end{Schunk}

Another interest of diagnosing is to examine how the index value
progresses for the points on the interpolation path since the projection
on these bases are played by the tour animation. Trace plots are created
with plotting the index value of the interpolation points against time.
Figure \ref{fig:toy-interp} presents the trace plot of the two
optimisation routines as \ref{fig:toy-search}. The trace plot is smooth
in both plots and this is a character of the holes index function used.
An interesting discovery is that with optimiser SAJO, the interpolator,
in iteration 6-8, first passes through some bases with higher index
value than the anchor bases.

@@

\caption[1D projection on the 5-variable dataset \code{boa5}5) with two optimisers]{1D projection on the 5-variable dataset \code{boa5}5) with two optimisers: Simulated Annealing (SA) and Pseudo Derivative (PD). All the bases in pseudo derivative has been flipped positive so that the two paths finish close to each other and to the theoretical best.}\label{fig:toy-pca}
\end{figure}
\end{Schunk}

Apart from checking the progression of an optimiser, looking at how the
points evaluated by the optimisers look like in the space is another
interest. Given the orthonormality constraint, the space of projection
bases \(\mathbf{A}_{p \times d}\) is a \(p \times d\) dimension sphere
and dimension reduction methods, i.e.~principal component analysis is
applied first to project all the bases onto a 2D space. In a projection
pursuit guided tour optimisation, there are various type of bases
involved: 1) The starting basis; 2) The anchor bases that have the
highest index value in each iteration; 3) The search bases that an
optimiser evaluated to produce the anchor bases; 4) The interpolating
bases on the interpolation path; and finally 5) the ending basis. The
importance of these basis differs with the most important ones being the
start, interpolating and ending bases. Anchor and search bases can be
turned on with argument \texttt{details\ =\ TRUE}.

Several annotations have been made to facilitate better reading of this
plot.
